---
title: "SKU Turnover Rate Analysis — 2023"
author: "Nandni Talreja"
format: html
editor: visual
---

## Overview

This file calculates SKU Turnover Rate for Amazon sales data covering January–September 2023.

**Turnover Rate = Total Units Sold / Average Inventory**

A higher turnover rate means inventory is selling and cycling quickly (good).
A lower rate means stock is sitting unsold (costly).

We calculate turnover across three time windows:
- January → June
- January → September
- June → September

---

```{r setup, message=FALSE, warning=FALSE}
library(readxl)
library(tidyverse)
library(lubridate)
library(kableExtra)
library(ggplot2)
library(writexl)
```

---

## 1. Load Data

```{r load-data, message=FALSE, warning=FALSE}
# Main Amazon order data (Jan–Sept 2023)
amazon_data <- read_xlsx(here::here("data", "amazon_jan_to_sept_2024.xlsx"))

# Inventory snapshots taken at three points in the year
# These come from Amazon Seller Central inventory reports
data_jan  <- read_xlsx(here::here("data", "inventory_jan.xlsx"))
data_june <- read_csv(here::here("data",  "inventory_june.csv"))
data_sept <- read_csv(here::here("data",  "inventory_sept.csv"))

# Quick check on dimensions
glimpse(amazon_data)
```

---

## 2. Parse Dates & Calculate Monthly Sales

```{r monthly-sales}
# Parse the purchase-date column to proper datetime
amazon_data$`purchase-date` <- ymd_hms(amazon_data$`purchase-date`)

# Extract year and month from the purchase date
amazon_data <- amazon_data %>%
  mutate(
    Year  = year(`purchase-date`),
    Month = month(`purchase-date`, label = TRUE, abbr = TRUE)
  ) %>%
  relocate(Year, Month, .after = `purchase-date`)

# Monthly sales aggregated by SKU
monthly_sales <- amazon_data %>%
  group_by(sku, Year, Month) %>%
  summarize(
    MonthlySales = sum(`item-price`, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(Year, Month)

# Yearly sales aggregated by SKU
yearly_sales <- amazon_data %>%
  group_by(sku, Year) %>%
  summarize(
    YearlySales = sum(`item-price`, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(Year)
```

---

## 3. Build Monthly Quantity Dataset

```{r monthly-quantity}
# Filter to 2023 data only
data_23 <- amazon_data %>%
  filter(Year == 2023) %>%
  select(sku, Month, quantity, `item-price`) %>%
  arrange(Month)

# Aggregate units sold per SKU for each month separately
# We do this individually so we can control the column names for the turnover formula
months_list <- list(
  jan  = "Jan", feb = "Feb", mar = "Mar",
  apr  = "Apr", may = "May", june = "Jun",
  july = "Jul", aug = "Aug", sept = "Sep"
)

monthly_qty <- map(months_list, function(m) {
  data_23 %>%
    filter(Month == m) %>%
    group_by(sku) %>%
    summarise(qty = sum(quantity, na.rm = TRUE), .groups = "drop")
})

# Rename qty columns to be month-specific before joining
monthly_qty <- imap(monthly_qty, function(df, name) {
  df %>% rename(!!paste0("quantity_", name) := qty)
})

# Join all months into one wide dataset by SKU
data_2023 <- reduce(monthly_qty, full_join, by = "sku")

# Calculate COGS per month (unit cost = $2.30)
UNIT_COST <- 2.30

data_2023 <- data_2023 %>%
  mutate(across(starts_with("quantity_"), 
                ~ . * UNIT_COST, 
                .names = "cogs_{.col}")) %>%
  rename_with(~ str_replace(., "cogs_quantity_", "cogs_"), 
              starts_with("cogs_quantity_"))
```

---

## 4. Calculate Turnover Rates

```{r turnover-calc}
# Pull inventory counts from each snapshot — rename for clarity
inv_jan  <- data_jan  %>% select(sku, available) %>% rename(available_jan  = available)
inv_june <- data_june %>% select(sku, available) %>% rename(available_june = available)
inv_sept <- data_sept %>% select(sku, available) %>% rename(available_sept = available)

# Join inventory snapshots with monthly sales data
turnover <- inv_jan %>%
  full_join(inv_june, by = "sku") %>%
  full_join(inv_sept, by = "sku") %>%
  full_join(data_2023, by = "sku")

# Replace NAs with 0 (SKUs not in a given snapshot had 0 inventory or 0 sales)
turnover[is.na(turnover)] <- 0

# --- Average Inventory for each window ---
# Formula: (Opening Inventory + Closing Inventory) / 2
avg_inv_jan_june <- (turnover$available_jan  + turnover$available_june) / 2
avg_inv_jan_sept <- (turnover$available_jan  + turnover$available_sept) / 2
avg_inv_june_sept <- (turnover$available_june + turnover$available_sept) / 2

# --- COGS per window (sum of monthly COGS) ---
cogs_jan_june <- turnover %>%
  group_by(sku) %>%
  summarise(cogs_jan_june = sum(cogs_jan, cogs_feb, cogs_mar, 
                                 cogs_apr, cogs_may, cogs_june, na.rm = TRUE))

cogs_jan_sept <- turnover %>%
  group_by(sku) %>%
  summarise(cogs_jan_sept = sum(cogs_jan, cogs_feb, cogs_mar, cogs_apr,
                                 cogs_may, cogs_june, cogs_july, 
                                 cogs_aug, cogs_sept, na.rm = TRUE))

cogs_june_sept <- turnover %>%
  group_by(sku) %>%
  summarise(cogs_june_sept = sum(cogs_june, cogs_july, 
                                  cogs_aug, cogs_sept, na.rm = TRUE))

# Join COGS back to turnover table
turnover <- turnover %>%
  full_join(cogs_jan_june,  by = "sku") %>%
  full_join(cogs_jan_sept,  by = "sku") %>%
  full_join(cogs_june_sept, by = "sku")

# --- Total units sold per window ---
turnover <- turnover %>%
  mutate(
    total_sales_jan_june = quantity_jan + quantity_feb + quantity_mar +
                           quantity_apr + quantity_may + quantity_june,
    total_sales_jan_sept = total_sales_jan_june + quantity_july + 
                           quantity_aug + quantity_sept,
    total_sales_june_sept = quantity_june + quantity_july + 
                            quantity_aug + quantity_sept
  )

# --- Avoid division by zero: if avg inventory is 0, set to 1 ---
turnover <- turnover %>%
  mutate(
    avg_inv_jan_june  = ifelse(avg_inv_jan_june  == 0, 1, avg_inv_jan_june),
    avg_inv_jan_sept  = ifelse(avg_inv_jan_sept  == 0, 1, avg_inv_jan_sept),
    avg_inv_june_sept = ifelse(avg_inv_june_sept == 0, 1, avg_inv_june_sept)
  )

# --- Calculate Turnover Rate for each window ---
turnover <- turnover %>%
  mutate(
    turnover_jan_june  = total_sales_jan_june  / avg_inv_jan_june,
    turnover_jan_sept  = total_sales_jan_sept  / avg_inv_jan_sept,
    turnover_june_sept = total_sales_june_sept / avg_inv_june_sept,
    # Rounded versions for reporting
    turnover_jan_june_r  = ceiling(turnover_jan_june),
    turnover_jan_sept_r  = ceiling(turnover_jan_sept),
    turnover_june_sept_r = ceiling(turnover_june_sept)
  )
```

---

## 5. Results Summary

```{r results-summary}
# Clean results table for reporting
sku_turnovers <- turnover %>%
  select(sku,
         cogs_jan_june,  avg_inv_jan_june,  turnover_jan_june,
         cogs_jan_sept,  avg_inv_jan_sept,  turnover_jan_sept,
         cogs_june_sept, avg_inv_june_sept, turnover_june_sept)

sku_turnovers %>%
  arrange(desc(turnover_jan_sept)) %>%
  head(20) %>%
  kable(digits = 2, caption = "Top 20 SKUs by Jan–Sept Turnover Rate") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

---

## 6. Flag Low-Performing SKUs

```{r flag-low-skus}
# Calculate average turnover across all three windows per SKU
turnover_mean <- turnover %>%
  rowwise() %>%
  mutate(avg_turnover = mean(c(turnover_jan_june, 
                                turnover_jan_sept, 
                                turnover_june_sept), 
                              na.rm = TRUE)) %>%
  ungroup()

# Threshold: SKUs with avg turnover < 3 are flagged as underperformers
# Most turnover values cluster between 0–10; < 3 = very slow moving
low_skus <- turnover_mean %>%
  filter(avg_turnover < 3) %>%
  select(sku, avg_turnover)

zero_skus <- turnover_mean %>%
  filter(avg_turnover == 0) %>%
  select(sku, avg_turnover)

cat("Total low-turnover SKUs (< 3):", nrow(low_skus), "\n")
cat("Zero-turnover SKUs (no sales at all):", nrow(zero_skus), "\n")
cat("Recommendation: Review zero-turnover SKUs for discontinuation\n")

# Show zero-turnover SKUs with their inventory levels
# These are SKUs we're paying to store but not selling
zero_skus %>%
  left_join(sku_turnovers, by = "sku") %>%
  arrange(desc(avg_inv_jan_sept)) %>%
  kable(digits = 2, caption = "Zero-Sales SKUs Still Holding Inventory") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

---

## 7. Export Results

```{r export}
# Attach product names for stakeholder-readable output
product_lookup <- amazon_data %>%
  select(sku, `product-name`) %>%
  distinct(sku, .keep_all = TRUE)

final_turnover <- turnover %>%
  select(sku, 
         turnover_jan_june, turnover_jan_sept, turnover_june_sept,
         turnover_jan_june_r, turnover_jan_sept_r, turnover_june_sept_r,
         total_sales_jan_sept) %>%
  left_join(product_lookup, by = "sku") %>%
  relocate(`product-name`, .after = sku) %>%
  arrange(desc(turnover_jan_sept))

# Write to Excel for stakeholder review
writexl::write_xlsx(
  list("Turnover Summary" = final_turnover,
       "Full Turnover Data" = turnover),
  here::here("outputs", "sku_turnover_results.xlsx")
)

cat("Results exported to outputs/sku_turnover_results.xlsx\n")
```
