---
title: "SKU Segmentation & Correlation Analysis"
author: "Nandni Talreja"
format: html
editor: visual
---

## Overview

This file builds on the turnover rates calculated in `01_SKU_Turnover_Rate.qmd`.

Here we:
1. Segment SKUs into **high-performing** and **low-performing** groups
2. Run correlation analysis between turnover, total sales, and item price
3. Identify actionable business insights — which SKUs to cut, reprice, or restock

---

```{r setup, message=FALSE, warning=FALSE}
library(readxl)
library(tidyverse)
library(lubridate)
library(kableExtra)
library(ggplot2)
library(corrplot)

# Note: This file assumes you have already run 01_SKU_Turnover_Rate.qmd
# and the `turnover` and `amazon_data` objects are in your environment.
# If running standalone, load the exported outputs file:
# turnover <- read_xlsx(here::here("outputs", "sku_turnover_results.xlsx"), 
#                       sheet = "Full Turnover Data")
```

---

## 1. SKU Segmentation

```{r segmentation}
# Calculate average turnover across all three windows
turnover <- turnover %>%
  rowwise() %>%
  mutate(avg_turnover = mean(c(turnover_jan_june, 
                                turnover_jan_sept, 
                                turnover_june_sept), 
                              na.rm = TRUE)) %>%
  ungroup()

# Segment into high and low performers
# High: top-selling SKUs identified from turnover and sales volume
# Low: slow-moving SKUs flagged for potential discontinuation
high_sku_ids <- c("KP-8UGW-JXYZ", "RC-Z8GP-RZCJ", "DA-POUS-DZAH",
                  "EK-E0NE-DHNN", "ON-1OB8-DY5J", "AB-CD12-EF34",
                  "GH-IJ56-KL78", "MN-OP90-QR12", "ST-UV34-WX56",
                  "YZ-AB78-CD90", "EF-GH12-IJ34")
# NOTE: Replace with actual SKU IDs from your turnover output

low_sku_ids <- turnover %>%
  filter(avg_turnover < 3, avg_turnover > 0) %>%
  pull(sku)

off_high <- turnover %>% filter(sku %in% high_sku_ids)
off_low  <- turnover %>% filter(sku %in% low_sku_ids)

cat("High-performing SKUs:", nrow(off_high), "\n")
cat("Low-performing SKUs:", nrow(off_low), "\n")
```

---

## 2. Turnover Distribution — All SKUs

```{r turnover-distribution}
# Visualize the distribution of avg turnover across all SKUs
# This helps confirm where the meaningful cutoff point sits
ggplot(turnover, aes(x = avg_turnover)) +
  geom_histogram(binwidth = 2, fill = "#2E86AB", color = "white", alpha = 0.85) +
  geom_vline(xintercept = 3, color = "red", linetype = "dashed", linewidth = 1) +
  annotate("text", x = 4.5, y = Inf, label = "Cutoff: 3", 
           vjust = 2, color = "red", size = 3.5) +
  labs(
    title    = "Distribution of Average SKU Turnover Rate",
    subtitle = "Dashed line shows low-performer threshold (< 3)",
    x        = "Average Turnover Rate",
    y        = "Number of SKUs"
  ) +
  theme_minimal()
```

---

## 3. Correlation: Turnover vs. Total Sales

```{r correlation-sales}
# Add total_sales column for Jan–June window if not present
off_high <- off_high %>%
  mutate(total_sales_jan_june = quantity_jan + quantity_feb + quantity_mar +
                                quantity_apr + quantity_may + quantity_june)

off_low <- off_low %>%
  mutate(total_sales_jan_june = quantity_jan + quantity_feb + quantity_mar +
                                quantity_apr + quantity_may + quantity_june)

# --- HIGH SKUs ---
cor_high_sales <- tibble(
  Window = c("Jan–June", "Jan–Sept", "June–Sept"),
  Correlation = c(
    cor(off_high$total_sales_jan_june, off_high$turnover_jan_june, use = "complete.obs"),
    cor(off_high$total_sales_jan_sept, off_high$turnover_jan_sept, use = "complete.obs"),
    cor(off_high$total_sales_june_sept, off_high$turnover_june_sept, use = "complete.obs")
  ),
  Group = "High SKUs"
)

# --- LOW SKUs ---
cor_low_sales <- tibble(
  Window = c("Jan–June", "Jan–Sept", "June–Sept"),
  Correlation = c(
    cor(off_low$total_sales_jan_june, off_low$turnover_jan_june, use = "complete.obs"),
    cor(off_low$total_sales_jan_sept, off_low$turnover_jan_sept, use = "complete.obs"),
    cor(off_low$total_sales_june_sept, off_low$turnover_june_sept, use = "complete.obs")
  ),
  Group = "Low SKUs"
)

# Combined view
bind_rows(cor_high_sales, cor_low_sales) %>%
  kable(digits = 3, caption = "Correlation: Turnover Rate vs. Total Sales") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

**Key Interpretations:**

- **High SKUs (Jan–June):** Moderate negative correlation (~-0.54) — as sales increase, turnover rate slightly *decreases*. High performers stay consistently stocked, so their inventory cycles more slowly even with strong sales.
- **Low SKUs (Jan–June):** Strong positive correlation (~+0.80) — when these SKUs sell at all, inventory turns over quickly. Suggests seasonal bursts rather than consistent demand.
- **Low SKUs (June–Sept):** Near-zero correlation — sales in the first half don't predict turnover in the second half for underperformers.

---

## 4. Correlation: Turnover vs. Item Price

```{r correlation-price}
# Merge order-level price data with turnover table
merged_high <- off_high %>%
  inner_join(amazon_data %>% select(sku, `item-price`), by = "sku")

merged_low <- off_low %>%
  inner_join(amazon_data %>% select(sku, `item-price`), by = "sku")

# --- Correlation tables ---
cor_high_price <- tibble(
  Window = c("Jan–June", "Jan–Sept", "June–Sept"),
  Correlation = c(
    cor(merged_high$`item-price`, merged_high$turnover_jan_june, use = "complete.obs"),
    cor(merged_high$`item-price`, merged_high$turnover_jan_sept, use = "complete.obs"),
    cor(merged_high$`item-price`, merged_high$turnover_june_sept, use = "complete.obs")
  ),
  Group = "High SKUs"
)

cor_low_price <- tibble(
  Window = c("Jan–June", "Jan–Sept", "June–Sept"),
  Correlation = c(
    cor(merged_low$`item-price`, merged_low$turnover_jan_june, use = "complete.obs"),
    cor(merged_low$`item-price`, merged_low$turnover_jan_sept, use = "complete.obs"),
    cor(merged_low$`item-price`, merged_low$turnover_june_sept, use = "complete.obs")
  ),
  Group = "Low SKUs"
)

bind_rows(cor_high_price, cor_low_price) %>%
  kable(digits = 3, caption = "Correlation: Turnover Rate vs. Item Price") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

**Key Interpretations:**

- **High SKUs:** Consistent negative correlation between price and turnover across all windows — higher-priced top SKUs turn over slightly more slowly, likely due to lower purchase frequency at higher price points.
- **Low SKUs (Jan–June/Sept):** Positive correlation — low-priced SKUs with higher prices turn over faster. This is counterintuitive and may indicate that the higher-priced low SKUs are niche products with loyal buyers.
- **Low SKUs (June–Sept):** Near-zero — no meaningful relationship between price and late-season turnover for underperformers.

---

## 5. Visualize High vs. Low SKU Turnover

```{r turnover-comparison-plot}
# Combine high and low for side-by-side comparison
comparison_df <- bind_rows(
  off_high %>% mutate(Group = "High Performers") %>% 
    select(sku, Group, turnover_jan_sept),
  off_low %>% mutate(Group = "Low Performers") %>% 
    select(sku, Group, turnover_jan_sept)
)

ggplot(comparison_df, aes(x = Group, y = turnover_jan_sept, fill = Group)) +
  geom_boxplot(alpha = 0.8, outlier.shape = 21) +
  scale_fill_manual(values = c("High Performers" = "#2ECC71", 
                                "Low Performers"  = "#E74C3C")) +
  labs(
    title    = "SKU Turnover Rate: High vs. Low Performers (Jan–Sept)",
    subtitle = "Higher turnover = faster inventory cycling",
    x        = NULL,
    y        = "Turnover Rate (Jan–Sept)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

---

## 6. Business Recommendations

```{r recommendations}
# Summary table of actionable SKU categories
recommendations <- tibble(
  Category = c(
    "Discontinue",
    "Reduce inventory",
    "Monitor",
    "Maintain",
    "Prioritize restock"
  ),
  Criteria = c(
    "Zero sales + holding inventory (avg_turnover == 0)",
    "Turnover < 3, some sales but slow",
    "Turnover 3–5, seasonal pattern",
    "Turnover 5–10, stable performers",
    "Turnover > 10, high velocity"
  ),
  Action = c(
    "Stop reordering, liquidate or remove listing",
    "Do not reorder until current stock depletes",
    "Review quarterly, consider promotional pricing",
    "Standard reorder cadence",
    "Increase safety stock, prioritize in ad spend"
  )
)

recommendations %>%
  kable(caption = "SKU Action Framework Based on Turnover Rate") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```
