---
title: "2024 SKU Turnover & Sales Forecasting"
author: "Nandni Talreja"
format: html
editor: visual
---

## Overview

This file extends the turnover analysis to 2024 data (January–September) and adds **predictive modeling** for daily sales forecasting.

Two models are built and compared:
1. **Linear Regression** — uses lagged sales and day-of-week as features
2. **ARIMA** — captures time-series trend and seasonality automatically

The goal is to improve restocking timing by predicting demand windows.

---

```{r setup, message=FALSE, warning=FALSE}
library(readxl)
library(tidyverse)
library(lubridate)
library(kableExtra)
library(ggplot2)
library(writexl)
library(forecast)
```

---

## 1. Load 2024 Data

```{r load-data, message=FALSE, warning=FALSE}
# Combined Amazon order data for Jan–Sept 2024
# Each row = one order line item
amazon_data <- read_xlsx(here::here("data", "amazon_jan_to_sept_2024.xlsx"))

# Inventory snapshots (from Amazon Seller Central)
data_jan  <- read_xlsx(here::here("data", "inventory_jan.xlsx"))
data_june <- read_csv(here::here("data",  "inventory_june.csv"))
data_sept <- read_csv(here::here("data",  "inventory_sept.csv"))

# Net proceeds data (revenue after Amazon fees)
net_proceeds <- read_xlsx(here::here("data", "net_proceeds.xlsx"))

# Product name lookup table (one row per SKU)
product_lookup <- amazon_data %>%
  select(sku, `product-name`) %>%
  distinct(sku, .keep_all = TRUE)

cat("Rows loaded:", nrow(amazon_data), "\n")
cat("Unique SKUs:", n_distinct(amazon_data$sku), "\n")
```

---

## 2. Aggregate Monthly Sales by SKU

```{r monthly-aggregation}
# List of monthly data sheets (loaded separately from the source file)
# In the original data, each month was a separate sheet in amazon_2024.xlsx
# Here we derive them from the combined file instead

amazon_data <- amazon_data %>%
  mutate(
    `purchase-date` = as.Date(`purchase-date`),
    Month = month(`purchase-date`, label = TRUE, abbr = TRUE),
    Year  = year(`purchase-date`)
  )

# Aggregate units sold per SKU per month
monthly_by_sku <- amazon_data %>%
  filter(Year == 2024) %>%
  group_by(sku, Month) %>%
  summarise(sales = sum(quantity, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = Month, values_from = sales, 
              names_prefix = "sales_", values_fill = 0)

# Ensure all 9 months are present even if some SKUs had zero sales
expected_cols <- paste0("sales_", c("Jan","Feb","Mar","Apr","May",
                                     "Jun","Jul","Aug","Sep"))
for (col in expected_cols) {
  if (!col %in% names(monthly_by_sku)) monthly_by_sku[[col]] <- 0
}

# Calculate COGS per month (unit cost = $2.30 per pair)
UNIT_COST <- 2.30

data_2024 <- monthly_by_sku %>%
  mutate(across(starts_with("sales_"), 
                ~ . * UNIT_COST, 
                .names = "cogs_{.col}")) %>%
  rename_with(~ str_replace(., "cogs_sales_", "cogs_"), starts_with("cogs_sales_"))

glimpse(data_2024)
```

---

## 3. Calculate 2024 Turnover Rates

```{r turnover-2024}
# Rename inventory columns for clarity
inv_jan  <- data_jan  %>% select(sku, available) %>% rename(inventory_jan  = available)
inv_june <- data_june %>% select(sku, available) %>% rename(inventory_june = available)
inv_sept <- data_sept %>% select(sku, available) %>% rename(inventory_sept = available)

# Join inventory snapshots and sales data
turnover <- inv_jan %>%
  full_join(inv_june,  by = "sku") %>%
  full_join(inv_sept,  by = "sku") %>%
  full_join(data_2024, by = "sku")

turnover[is.na(turnover)] <- 0

# Average inventory for each window
avg_inv_jan_june  <- (turnover$inventory_jan  + turnover$inventory_june) / 2
avg_inv_jan_sept  <- (turnover$inventory_jan  + turnover$inventory_sept) / 2
avg_inv_june_sept <- (turnover$inventory_june + turnover$inventory_sept) / 2

# COGS per window
cogs_jan_june <- turnover %>%
  rowwise() %>%
  mutate(cogs_jan_june = sum(c_across(c(cogs_Jan, cogs_Feb, cogs_Mar, 
                                         cogs_Apr, cogs_May, cogs_Jun)))) %>%
  select(sku, cogs_jan_june) %>% ungroup()

cogs_jan_sept <- turnover %>%
  rowwise() %>%
  mutate(cogs_jan_sept = sum(c_across(starts_with("cogs_")))) %>%
  select(sku, cogs_jan_sept) %>% ungroup()

cogs_june_sept <- turnover %>%
  rowwise() %>%
  mutate(cogs_june_sept = sum(c_across(c(cogs_Jun, cogs_Jul, 
                                          cogs_Aug, cogs_Sep)))) %>%
  select(sku, cogs_june_sept) %>% ungroup()

turnover <- turnover %>%
  full_join(cogs_jan_june,  by = "sku") %>%
  full_join(cogs_jan_sept,  by = "sku") %>%
  full_join(cogs_june_sept, by = "sku")

# Total sales per window
turnover <- turnover %>%
  mutate(
    total_sales_jan_june  = sales_Jan + sales_Feb + sales_Mar + 
                            sales_Apr + sales_May + sales_Jun,
    total_sales_jan_sept  = total_sales_jan_june + 
                            sales_Jul + sales_Aug + sales_Sep,
    total_sales_june_sept = sales_Jun + sales_Jul + sales_Aug + sales_Sep
  )

# Guard against divide-by-zero
turnover <- turnover %>%
  mutate(
    avg_inv_jan_june  = ifelse(avg_inv_jan_june  == 0, 1, avg_inv_jan_june),
    avg_inv_jan_sept  = ifelse(avg_inv_jan_sept  == 0, 1, avg_inv_jan_sept),
    avg_inv_june_sept = ifelse(avg_inv_june_sept == 0, 1, avg_inv_june_sept),
    # Overall average inventory across all three windows
    avg_inventory = (avg_inv_jan_june + avg_inv_jan_sept + avg_inv_june_sept) / 3
  )

# Calculate turnover rates
turnover <- turnover %>%
  mutate(
    turnover_jan_june  = total_sales_jan_june  / avg_inv_jan_june,
    turnover_jan_sept  = total_sales_jan_sept  / avg_inv_jan_sept,
    turnover_june_sept = total_sales_june_sept / avg_inv_june_sept
  )

# Attach net proceeds (revenue after Amazon fees)
net_proceeds <- net_proceeds %>% rename(sku = SKU)

turnover_final <- turnover %>%
  full_join(net_proceeds,  by = "sku") %>%
  left_join(product_lookup, by = "sku") %>%
  relocate(`product-name`, .after = sku)

# Export for stakeholders
writexl::write_xlsx(
  list("Turnover 2024" = turnover_final),
  here::here("outputs", "sku_turnover_2024.xlsx")
)

cat("2024 turnover analysis exported.\n")
```

---

## 4. Daily Sales Trend

```{r daily-sales-plot}
# Aggregate to daily total sales across all SKUs
daily_sales <- amazon_data %>%
  filter(Year == 2024) %>%
  group_by(`purchase-date`) %>%
  summarize(total_sales = sum(quantity, na.rm = TRUE), .groups = "drop") %>%
  arrange(`purchase-date`)

# Add time features for modeling
daily_sales <- daily_sales %>%
  mutate(
    day_of_week = weekdays(`purchase-date`),
    month       = format(`purchase-date`, "%m")
  )

# Plot overall sales trend
ggplot(daily_sales, aes(x = `purchase-date`, y = total_sales)) +
  geom_line(color = "#2E86AB", alpha = 0.8) +
  geom_smooth(method = "loess", se = FALSE, color = "#E74C3C", 
              linetype = "dashed", linewidth = 0.8) +
  labs(
    title    = "Daily Amazon Sales — Jan to Sept 2024",
    subtitle = "Red dashed line = smoothed trend",
    x        = "Date",
    y        = "Units Sold"
  ) +
  theme_minimal()
```

---

## 5. Linear Regression Sales Forecast

```{r linear-model}
# Feature engineering: lagged sales capture recent momentum
# sales_lag_1 = yesterday's sales (short-term momentum)
# sales_lag_7 = same day last week (weekly seasonality)
daily_sales <- daily_sales %>%
  mutate(
    sales_lag_1 = lag(total_sales, 1),
    sales_lag_7 = lag(total_sales, 7)
  ) %>%
  drop_na()  # Remove rows with NAs from lag creation

# 80/20 train/test split (chronological — important for time series)
set.seed(123)
train_size <- floor(0.8 * nrow(daily_sales))
train_data <- daily_sales[1:train_size, ]
test_data  <- daily_sales[(train_size + 1):nrow(daily_sales), ]

cat("Training rows:", nrow(train_data), "\n")
cat("Test rows:", nrow(test_data), "\n")

# Fit linear regression model
lm_model <- lm(total_sales ~ sales_lag_1 + sales_lag_7 + day_of_week, 
               data = train_data)

summary(lm_model)
```

```{r lm-evaluation}
# Generate predictions on holdout test set
lm_predictions <- predict(lm_model, test_data)

# Evaluation metrics
mae_lm  <- mean(abs(lm_predictions - test_data$total_sales))
rmse_lm <- sqrt(mean((lm_predictions - test_data$total_sales)^2))

cat("Linear Regression — Test Set Performance:\n")
cat("  MAE  (Mean Absolute Error):", round(mae_lm, 2), "units\n")
cat("  RMSE (Root Mean Sq Error): ", round(rmse_lm, 2), "units\n")

# Actual vs Predicted plot
ggplot() +
  geom_line(data = test_data, 
            aes(x = `purchase-date`, y = total_sales, color = "Actual")) +
  geom_line(aes(x = test_data$`purchase-date`, y = lm_predictions, 
                color = "Predicted (LM)")) +
  scale_color_manual(values = c("Actual" = "#2E86AB", 
                                 "Predicted (LM)" = "#E74C3C")) +
  labs(
    title    = "Linear Regression: Actual vs. Predicted Daily Sales",
    subtitle = paste0("Test period | MAE: ", round(mae_lm, 1), 
                      " units | RMSE: ", round(rmse_lm, 1), " units"),
    x        = "Date",
    y        = "Units Sold",
    color    = NULL
  ) +
  theme_minimal()
```

---

## 6. ARIMA Time-Series Forecast

```{r arima-model}
# ARIMA (AutoRegressive Integrated Moving Average)
# auto.arima() selects the best p,d,q parameters automatically
# This model captures trend + seasonality without manual feature engineering

arima_model <- auto.arima(train_data$total_sales)
summary(arima_model)
```

```{r arima-forecast}
# Forecast for the length of the test set
arima_forecast <- forecast(arima_model, h = nrow(test_data))

# Evaluation metrics
mae_arima  <- mean(abs(arima_forecast$mean - test_data$total_sales))
rmse_arima <- sqrt(mean((arima_forecast$mean - test_data$total_sales)^2))

cat("ARIMA — Test Set Performance:\n")
cat("  MAE  (Mean Absolute Error):", round(mae_arima, 2), "units\n")
cat("  RMSE (Root Mean Sq Error): ", round(rmse_arima, 2), "units\n")

# Actual vs ARIMA predicted plot
ggplot() +
  geom_line(data = test_data,
            aes(x = `purchase-date`, y = total_sales, color = "Actual")) +
  geom_line(aes(x = test_data$`purchase-date`, 
                y = as.numeric(arima_forecast$mean),
                color = "Predicted (ARIMA)")) +
  scale_color_manual(values = c("Actual" = "#2E86AB", 
                                 "Predicted (ARIMA)" = "#27AE60")) +
  labs(
    title    = "ARIMA: Actual vs. Predicted Daily Sales",
    subtitle = paste0("Test period | MAE: ", round(mae_arima, 1), 
                      " units | RMSE: ", round(rmse_arima, 1), " units"),
    x        = "Date",
    y        = "Units Sold",
    color    = NULL
  ) +
  theme_minimal()
```

---

## 7. Model Comparison

```{r model-comparison}
# Side-by-side comparison of both models
model_comparison <- tibble(
  Model   = c("Linear Regression", "ARIMA"),
  MAE     = c(round(mae_lm, 2),    round(mae_arima, 2)),
  RMSE    = c(round(rmse_lm, 2),   round(rmse_arima, 2)),
  Notes   = c(
    "Uses lag features + day-of-week; interpretable",
    "Auto-selects time-series structure; no feature engineering needed"
  )
)

model_comparison %>%
  kable(caption = "Forecast Model Comparison — Test Set") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

**Business Implication:** Lower MAE = closer predictions = better restocking timing. Use the better-performing model to forecast demand 2–4 weeks ahead and trigger reorder alerts when predicted demand exceeds current safety stock levels.
